---
title: "Using roogledocs for presentations"
author: "Rob Challen"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using roogledocs for presentations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
resource_files:
  - docs/articles/example-template-slides.pdf
  - docs/articles/slides-after-update.pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE
)

suppressWarnings(library(tidyverse))
suppressWarnings(library(huxtable))
# devtools::load_all("~/Git/roogledocs/")
here::i_am("vignettes/web-only/roogleslides.Rmd")
```

## Initialising the library

Prior to doing any analysis we may have some form of template. This might be a
document skeleton or report template. It can contain empty tables, place-holder
images, and double-brace tags, all of which can be replaced by calculated content from
R. To do this we initialise the `roogledocs` library. 

```{r}

# There is a global flag to disable `roogledocs` in case you want to develop and
# test offline.
options('roogledocs.disabled'=FALSE)

# roogledocs stores an authentication token on your local hard drive.
options("roogledocs.tokenDirectory"="~/.roogledocs-test")
```

Most of the time you will be creating or updating a single document. For this
vignettes sake it is useful to be able to delete previous versions. The point of
`roogledocs` though is actually to work with a continuously updated document and
therefore deleting documents is usually not what you want to do. Likewise for
this vignette it is useful to get a copy of the Google doc as a PDF from R. This
may not be that useful in real life. The main function here though is the
`findOrCloneTemplate()` method which lets you find a Google doc by name, or
clone a template document if you can't find it. There are also equivalent
methods to find or create blank documents, or just find Google docs by name or
sharing URL if they already exist.

```{r}
roogledocs::delete_slides("roogleslides-demo",areYouSure = TRUE)
```

Sometimes (particularly if the `roogledocs` library has been updated) we get a 
`TokenResponseException`, saying the token has been expired or revoked. In this 
event explicitly re-authenticating the library can be done though a call to 
`roogledocs::reauth()`.

```{r eval=FALSE}
roogledocs::reauth()
```

Once authentication is working getting a 

```{r}
doc = roogledocs::slides_from_template(
  "roogleslides-demo",
  "https://docs.google.com/presentation/d/18jqzzDI1zBruO3Rc0RlzX_rxhsDiOuKbTM4vnwuimn4/edit?usp=sharing")
fs::dir_create(here::here("docs/articles/web-only"))
doc$saveAsPdf(here::here("docs/articles/web-only/example-template-slides.pdf"))
```

Running the chunk above should authenticate you and grab a publicly shared
template I created, and make a copy of it in your Google drive under the name
"roogledocs-demo". [The document template can be seen here](example-template-slides.pdf), or as the 
[original Google doc](https://docs.Google.com/document/d/1XnrBgBJFz7jEMYtw3o3YKbOuMdWvUkzIul4hb2B-SC4/edit?usp=sharing).

# Tabular data

Inserting tables in this document is done by index. There is already a blank
table 1 in the document. At the moment we support only `huxtable` tables and plain 
data-frames. The
following chunk creates a sample `huxtable` from the diamonds data set, applies
some formatting and replaces the content of table one in the template with this
data. The formatting is preserved more or less. There is only support for basic
text formatting, borders (black solid only at present), background colour, and
alignment. The table will respect column widths as a relative measure and the
command takes a overall table width parameter. Layout will then depend on the
content. Custom row heights are not supported.

```{r table1}
hux = diamonds %>% mutate(colorCat = ifelse(color <= "G", "D-G","G-J")) %>% group_by(cut,colorCat) %>% summarise(
  `Size (mean + sd)` = sprintf("%1.2f \u00B1 %1.2f",mean(carat),sd(carat)),
  `Cost (mean + sd)` = sprintf("%1.0f \u00B1 %1.0f",mean(price),sd(price))
) %>% huxtable::as_hux() %>%
  huxtable::theme_article() %>% 
  huxtable::set_all_padding(value = 0) %>%
  huxtable::merge_repeated_rows()

hux %>% roogledocs::as.long_format_table() %>% doc$updateTaggedTable(tagName = "diamonds-tag" )

hux
```

# Updating figures

A similar process exists for figures. We need to have the figure as a PNG image on the
local computer as a result of, for example, a ggplot. Once a local PNG is
available, it is temporarily uploaded to your Google drive, added to the
document and then temporary drive file deleted. In this example we update figure
1 replacing the `{{figure_1}}` tag in the original Google doc with the image:

```{r fig1}

dim = doc$slideDimensions()

g = ggplot(diamonds, aes(x=carat,y=price, colour=color))+geom_point()
figure_1 = roogledocs::ggplot_to_png(g, width=dim$width, height=dim$height)

# If the first parameter is passed as a variable and no tag is given, 
# as in this example the variable name is used as the tag and the image is 
# inserted into the document at the location of the tag:

doc$updateTaggedImage(figure_1)

# This is equivalent to:
# doc$updateTaggedImage(figure_1, tagName = "figure_1")


```

Updating a second figure can happen in the same way. The dimensions of the
image in the Google slides... resized to fit the container. If the container
is rotated the result will be rotated.

```{r fig2}
g2 = ggplot(diamonds, aes(x=cut,y=price, fill=cut))+
  geom_violin(draw_quantiles = c(0.95,0.5,0.05))+
  scale_fill_brewer()
figure_2 = roogledocs::ggplot_to_png(g2, width=5, height=4)

doc$updateTaggedImage(figure_2)
```


# Updating tagged text

If you want to update small textual results - e.g. results in the abstract of a
paper (similar to RMarkdown in-line chunks) you can place a double-brace tag
into the Google doc and replace this with text generated in R. The result is
inserted in the Google doc as a URL link so that further changes or updates in
code can find the tagged text. Links like this can be moved around the document,
or copied and pasted without losing the tag. You can get a
list of the tags present in a document like this:

```{r tags}
doc$tagsDefined()
```

Here we have 2 tags. The tags can be then set to specific content like this:

```{r autotext}
format(Sys.Date(),"%d/%m/%Y") %>% doc$updateTaggedText(tagName = "date")
diamonds_mean_sd = sprintf("%1.1f \u00B1 %1.1f",mean(diamonds$price),sd(diamonds$price)) 

# if we don't give a specific tag name then the variable name is used:
doc$updateTaggedText(diamonds_mean_sd)
```


# New content

Appending new content is also possible either as a simple styled text string,
with consistent formatting, or as a continuous block (or blocks) with different
styles, as specified in a data-frame. At the moment this is only possible at the
end of the document and is really designed if a document is being generated
completely from scratch. 

```{r new-text}

content = tibble::tribble(
  ~label, ~link, ~fontName, ~fontFace,
  "Roogledocs", "https://terminological.github.io/roogledocs/r-library/docs/", "Courier New", "plain",
  " is also able to add text at the end of the document with complex formatting. ", NA, NA, "plain",
  "Supporting fonts and font face formatting such as ",  NA, NA, "plain",
  "bold, ", NA, NA, "bold",
  "italic ", NA, NA, "italic",
  "and underlined", NA, NA, "underlined",
  " amongst other things.", NA, NA, "plain"
  )

doc$appendFormattedSlide("Adding new content", content)

content
```

It could be possible to combine writing new content and updating tagged text in
the same script to programmatically generate replacement content. Likewise this
could be used for captions of tables and figures when they are added. When you
write new content you can write in double-brace tags and these can then be
updated at a later stage for example.

# Citations

Assuming we have a bibliography file we can provide this to `roogledocs` and 
it will insert citations.

```{r}

doc$updateCitations(here::here("vignettes/web-only/test.bib"), citationStyle = "journal-of-infection")

```

Finally we can write out the new document to a PDF, mostly so we can see what we
have done. When we write out a document any `roogledocs` links are removed:

```{r}
doc$saveAsPdf(here::here("docs/articles/web-only/slides-after-update.pdf"))
```

After the analysis has run we have a new version of the Google slides which
should [look like this](slides-after-update.pdf).






