---
title: "Using roogledocs for documents"
author: "Rob Challen"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using roogledocs for documents}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
resource_files:
  - docs/articles/example-template-doc.pdf
  - docs/articles/example-after-update.pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE
)

suppressWarnings(library(tidyverse))
suppressWarnings(library(huxtable))
# devtools::load_all("~/Git/roogledocs/")
here::i_am("vignettes/web-only/roogledocs.Rmd")
```


```{r eval=FALSE}
remove.packages("roogledocs")
devtools::install_local("~/Git/roogledocs",upgrade = FALSE)
```

## Initialising the library

Prior to doing any analysis we may have some form of template. This might be a
document skeleton or report template. It can contain empty tables, place-holder
images, and double-brace tags, all of which can be replaced by calculated content from
R. To do this we initialise the `roogledocs` library. 

```{r}

# There is a global flag to disable `roogledocs` in case you want to develop and
# test offline.
options('roogledocs.disabled'=FALSE)

# roogledocs stores an authentication token on your local hard drive.
options("roogledocs.tokenDirectory"="~/.roogledocs-test")
```


```{r}
J = roogledocs::JavaApi$get(logLevel = "WARN")
x = J$RoogleDocs$new()
```

Most of the time you will be creating or updating a single document. For this
vignettes sake it is useful to be able to delete previous versions. The point of
`roogledocs` though is actually to work with a continuously updated document and
therefore deleting documents is usually not what you want to do. Likewise for
this vignette it is useful to get a copy of the Google doc as a PDF from R. This
may not be that useful in real life. The main function here though is the
`findOrCloneTemplate()` method which lets you find a Google doc by name, or
clone a template document if you can't find it. There are also equivalent
methods to find or create blank documents, or just find Google docs by name or
sharing URL if they already exist.

```{r}
roogledocs::delete_document("roogledocs-demo",areYouSure = TRUE)
```

Sometimes (particularly if the `roogledocs` library has been updated) we get a 
`TokenResponseException`, saying the token has been expired or revoked. In this 
event explicitly re-authenticating the library can be done though a call to 
`roogledocs::reauth()`.

```{r eval=FALSE}
roogledocs::reauth()
```

Once authentication is working we get a new document based on a template I 
creates and shared:

```{r}
doc = roogledocs::doc_from_template(
  "roogledocs-demo",
  "https://docs.Google.com/document/d/1XnrBgBJFz7jEMYtw3o3YKbOuMdWvUkzIul4hb2B-SC4/edit?usp=sharing"
)
fs::dir_create(here::here("docs/articles/web-only"))
doc$saveAsPdf(here::here("docs/articles/web-only/example-template-doc.pdf"))
```

Running the chunk above should authenticate you and grab a publicly shared
template I created, and make a copy of it in your Google drive under the name
"roogledocs-demo". [The document template can be seen here](example-template-doc.pdf), or as the 
[original Google doc](https://docs.Google.com/document/d/1XnrBgBJFz7jEMYtw3o3YKbOuMdWvUkzIul4hb2B-SC4/edit?usp=sharing).

# Tabular data

Inserting tables in this document is done by index. There is already a blank
table 1 in the document. At the moment we support only `huxtable` tables and plain 
data-frames. The
following chunk creates a sample `huxtable` from the diamonds data set, applies
some formatting and replaces the content of table one in the template with this
data. The formatting is preserved more or less. There is only support for basic
text formatting, borders (black solid only at present), background colour, and
alignment. The table will respect column widths as a relative measure and the
command takes a overall table width parameter. Layout will then depend on the
content. Custom row heights are not supported.

```{r table1}
hux = diamonds %>% mutate(colorCat = ifelse(color <= "G", "D-G","G-J")) %>% group_by(cut,colorCat) %>% summarise(
  `Size (mean + sd)` = sprintf("%1.2f \u00B1 %1.2f",mean(carat),sd(carat)),
  `Cost (mean + sd)` = sprintf("%1.0f \u00B1 %1.0f",mean(price),sd(price))
) %>% huxtable::as_hux() %>%
  huxtable::theme_article() %>% 
  huxtable::set_all_padding(value = 0) %>%
  huxtable::merge_repeated_rows()

table_1 = hux %>% roogledocs::as.long_format_table() 
doc$updateTaggedTable(table_1, tableWidthInches = 4)

hux
```

# Updating figures

A similar process exists for figures. We need to have the figure as a PNG image on the
local computer as a result of, for example, a ggplot. Once a local PNG is
available, it is temporarily uploaded to your Google drive, added to the
document and then temporary drive file deleted. In this example we update figure
1 replacing the `{{figure_1}}` tag in the original Google doc with the image:

```{r fig1}
g = ggplot(diamonds, aes(x=carat,y=price, colour=color))+geom_point()
figure_1 = roogledocs::ggplot_to_png(g, width=6, height=3)

# If the first parameter is passed as a variable and no tag is given, 
# as in this example the variable name is used as the tag and the image is 
# inserted into the document at the location of the tag:

doc$updateTaggedImage(figure_1)

# This is equivalent to:
# doc$updateTaggedImage(figure_1, tagName = "figure_1")

g
```

Updating a second figure can happen in the same way, but in this case we use the
alternative by specifying the index of insertion. The dimensions of the
image in the Google doc should exactly match the dimensions of the PNG file
saved from R. This means if you change the size of an image in R it will be
changed in the document. Image dimensions are therefore important to decide on
in R. If the figure or table had not already existed in the target Google doc
(e.g. because you started with a blank document) they would simply have been
uploaded and added at the end of the document as a sequentially numbered image.
If you rearrange the order figures in the Google doc it is up to you to fix the
indexes in your code. Captions are not handled here at all as it is assumed that
the captions will be maintained in the Google doc and not in R (see "Updating
tagged text" section later).

```{r fig2}
g2 = ggplot(diamonds, aes(x=cut,y=price, fill=cut))+
  geom_violin(draw_quantiles = c(0.95,0.5,0.05))+
  scale_fill_brewer()+
  theme(axis.text.x = element_text(angle = 15, vjust=1,hjust=1))
filename = roogledocs::ggplot_to_png(g2, width=4, height=3)

# The figure index has to be calculated with respect to any changes that have 
# already been made in the document. In this case inserting figure 1 before
# means this is inserted in the right place, but it is up to the user to make 
# sure this is right.
filename %>% doc$updateFigure(figureIndex = 2)
g2
```

Along with updating a figure within a document as a png it is useful to also be
able to keep a copy of the figure with the document in a publication ready
format such as pdf. We might also be generating supplementary material / or
data, and or tables in a separate word document. In my work flow all of these
are generated by R scripts. Automatically uploading these documents to a folder in Google drive
makes managing the output of an analysis fairly straightforward. Because Google Drive
can have multiple files of the same name in the same directory the behaviour 
must be specified if the file already exists.

```{r}
doc$uploadSupplementaryFiles(absoluteFilePath = figure_1,overwrite = TRUE)
```

# Updating tagged text

If you want to update small textual results - e.g. results in the abstract of a
paper (similar to RMarkdown in-line chunks) you can place a double-brace tag
into the Google doc and replace this with text generated in R. The result is
inserted in the Google doc as a URL link so that further changes or updates in
code can find the tagged text. Links like this can be moved around the document,
or copied and pasted without losing the tag. You can get a
list of the tags present in a document like this:

```{r tags}
doc$tagsDefined()
```

Here we have 2 tags. The tags can be then set to specific content like this:

```{r autotext}
format(Sys.Date(),"%d/%m/%Y") %>% doc$updateTaggedText(tagName = "table_1_update_date")
diamonds_mean_sd = sprintf("%1.1f \u00B1 %1.1f",mean(diamonds$price),sd(diamonds$price)) 

# if we don't give a specific tag name then the variable name is used:
doc$updateTaggedText(diamonds_mean_sd)
```

If the tags have become broken, you may
need to revert all the tags in a document so you can easily see where the good
tags are (or to identify if some tags have been lost by copy paste). This can be
done with the `doc$revertTags()` function, which puts the double-brace tags back
in the document removing the auto-text. It is possible to use this kind of
approach to automatically add in captions for figures or tables. 

# New content

Appending new content is also possible either as a simple styled text string,
with consistent formatting, or as a continuous block (or blocks) with different
styles, as specified in a data-frame. At the moment this is only possible at the
end of the document and is really designed if a document is being generated
completely from scratch. A minimal subset of text formatting is supported
although by using named styles you can do more. It is not actually that hard to
support more styles but at the moment this is not the primary use case for 
`roogledocs`.

```{r new-text}

doc$appendText("\nAdding new content\n","HEADING_1")

content = tibble::tribble(
  ~label, ~link, ~fontName, ~fontFace,
  "Roogledocs", "https://terminological.github.io/roogledocs/r-library/docs/", "Courier New", "plain",
  " is also able to add text at the end of the document with complex formatting. ", NA, NA, "plain",
  "Supporting fonts and font face formatting such as ",  NA, NA, "plain",
  "bold, ", NA, NA, "bold",
  "italic ", NA, NA, "italic",
  "and underlined", NA, NA, "underlined",
  " amongst other things.\n\n", NA, NA, "plain"
  )

doc$appendFormattedParagraph(content)

content
```

It could be possible to combine writing new content and updating tagged text in
the same script to programmatically generate replacement content. Likewise this
could be used for captions of tables and figures when they are added. When you
write new content you can write in double-brace tags and these can then be
updated at a later stage for example.

Citations are referenced as a tag of the format `{{cite:XXXX;YYYY}}` where X and Y 
are bibtext ids. References will be updated if present or if not  

```{r}

doc$updateCitations(here::here("vignettes/web-only/test.bib"), citationStyle = "journal-of-infection")

```

Finally we can write out the new document to a PDF, mostly so we can see what we
have done. When we write out a document any `roogledocs` links are removed:

```{r}
doc$saveAsPdf(here::here("docs/articles/web-only/example-after-update.pdf"))
```

After the analysis has run we have a new version of the Google document which
should [look like this](example-after-update.pdf).

# Limitations and further options

There are a lot of possible ways to extend `roogledocs`. The current 
implementation is in evolution to meet my own needs. 

* For example, additional formatting options such as text colour is not
implemented but would be relatively straightforward. 
* Currently there is no support for lists, which should be fairly
simple, but I didn't have a need yet.
* Absolutely positioned images are ignored completely. This is probably a good
thing as it lets you have a logo within the document for example without messing
up the dynamic images from R.
* Building a Google docs based drop in for the `officer` library for MS word
could be good if it means we can leverage that interface for other uses (e.g.
knitr support). 





